---
title: "Texas Covid Time Series PLots"
author: "The Covid Project and Craig Slinkman"
date: "7/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Texas Covid Data Plots

This RMarkdown document plots the time series data provided by 
[The Covid Project](https://covidtracking.com).

```{r load_packages}
require(tidyverse)                          # I live in the tidyverse ...
require(lubridate)                          # For advanced date processing ...
require( cowplot )                          # For academic plotting ...
require(formattable)                        # For fixed decimal formatting ...
```
 
## Load required_functions

```{r load_required functions}
source('D:/R-Projects/TexasCovid/functions/get_texas_daily_covid_data.R')
source('D:/R-Projects/TexasCovid/functions/plot_ts.R')
 
```

## Set cowplot as the default theme


The default theme in ggplot produces a high ratio of ink to data points.  Therefore, I prefer to use the cowplot theme.  We loaded the theme in load packages chunk but we need to set itn here.

```{r set_cowpllot_as_default}

# Note: As of version 1.0.0, cowplot does not change the default ggplot2 theme
# anymore. To recover the previous behavior, execute:
# theme_set(theme_cowplot())

theme_set(theme_cowplot())
```
 
## Get the Texas Covid data

We read the data from the Covid Project data.

```{r get_daily_covid_data}

###############################################################################
# We get the covid data using the functions/get_texas_daily_covid_data.
###############################################################################

texas <- get_texas_daily_covid_data()
```

## Graphical Analysis of Tests

This section reports the analysis of the Texas testing results.

### Positive tests

#### Definition of a positive test

Total number of people with confirmed OR probable COVID-19 reported by the state or territory (per the expanded CSTE case definition of April 5th, 2020 approved by the CDC). A confirmed case is a person who has a positive test result from an FDA approved diagnostic molecular test. A probable case is a person who either has presentable symptoms WITH epidemiological evidence or has BOTH a positive presumptive laboratory test AND also EITHER presentable symptoms OR epidemiological evidence. Epidemiological evidence refers either to close-proximity contact with a known case or travel history to an area with high disease incidence. According to the guidelines, FDA approved antibody and antigen tests are considered part of presumptive laboratory evidence.

#### Positive plot

```{r plot_positive_cases}

#####################################################################
# We compute y-axis tick marks.  Noe that the positive case variable is in the
# third column.
#####################################################################

variable <- "positive"
variable_name <- "Cummulative positive tests"
title   <- "Texas nCovid-19 positive tests"

plot_ts( texas,
         variable,
         variable_name,
         title )
   
```

### Negative test cases

#### Definition of negative test

Individuals with a completed viral test that returned a negative result. For states / territories that do not report this number directly, we compute it using one of several methods, depending on which data points the state provides.  

#### Time series plot of negative tests

```{r plot)negative_cases}

variable       <- "negative"
variable_label <- "Cummulative negative tests"
title          <- "Texas nCovid-19 negative tests"
 
plot_ts( texas,
         variable,
         variable_name,
         title )
```

