---
title: "texas_county_covid_data"
author: "The Covid Project and Craig Slinkman"
date: "8/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This document shows how to access the county covid deaths provided by the New
York Times via their [GitHub Covid-site](https://github.com/nytimes/covid-19-data).

We start by loading the R packages and written functions we need to produce this
document.

```{r load_pacakges}

require(tidyverse)                          # I live in the tidyverse ....
require(lubridate)                          # For advanced date processing ...
require( stringr )                          # For string processing
```

Now we load the required functions.

```{r load_functions}

source('D:/R-Projects/TexasCovid/functions/get_us_county_cases_and_deaths.R')
source('D:/R-Projects/TexasCovid/functions/get_us_counties_population.R')
source('D:/R-Projects/TexasCovid/functions/get_texas_fips_codes.R')

```


## Reading the United Statescovid data
function
We access the United States Covid data below:

```{r read_county_covid_data}

###############################################################################
# We read the US county data case and deaths data and store the data in the
# US tibble.
###############################################################################

us_covid <- get_us_county_cases_and_deaths()      # Read NY Times covid data.
us_covid                                          # Display tibble
```

# We extract the Texas state data from the **us_coivid" tibble

```{r extract_data}
tx_covid <- 
    us_covid %>%
        filter( state == "Texas" )
tx_covid
```

## Load the United States Census data

```{r get_us_census_dara}
us_census <- get_us_counties_population()
us_census
```

## Extract Texas census data from US census data

```{r extract_texas}
tx_census <- 
    us_census %>% 
        filter( STNAME == "Texas" )
tx_census
```

## Keep oonly county rows 

The file design of the census data contains nore than one l3evel of goverment.  We only want to keep the county data.  So well extract the texas county data

```{r extract_counties}
n <- dim( tx_census)[1]
tx_counties <- tx_census[2:n,]
tx_counties
```

## Keep only required columns 

```{r keep_required_data}
tx_counties <- 
    tx_counties %>% 
        select( CTYNAME,
                CENSUS2010POP,
                POPESTIMATE2019 )
tx_counties
```

## Rename columns
We rename the columns so that follow Hadley Wickham's tidyverse naming cponventiaon that I use.

```{r rename_columns}
tx_counties <- 
    tx_counties %>% 
        rename( "county"                    = "CTYNAME",
                "census_population_2010"    = "CENSUS2010POP",
                "estimated_population_2019" = "POPESTIMATE2019" )
tx_counties   
```

## Estimat population for 2019
We estimate the 2020 population for each county by computing the 9 year grown that and multiplying that by the estimated 2019 population.

```{r Estimate_2020_populatoion}
tx_counties <-  
    tx_counties %>% 
        mutate( growth <- ( estimated_population_2019 / 
                            census_population_2010 )^(1/9),
                estimated_population_2020 <- 
                    growth * estimated_population_2019 )
tx_counties
```

## Join the covid tibble to the tx_county tibble ny count name

```{r join_covid_to_tx_counties}
tx_covid <- 
    left_join( tx_covid,
               tx_counties )
tx_covid
```

